### 2022-03-22 outputs/paper/paper.Rmd
added comment

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Lantao Cui
  - Huakun Shen
  - Sijian Wu
thanks: "Code and data are available at: https://github.com/HuakunShen/Discrimination-Analysis-Canadian-2014-GSS"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)

library(tidyverse)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(here)
library(gridExtra)

source(here::here("scripts/constants.R"))
source(here::here("scripts/cross-compare-barplot.R"))
source(here::here("scripts/cross-compare-plots.R"))
source(here::here("scripts/crime-count.R"))
```

# Introduction

Discrimination had always been a worldwide issue, in the past, people used to judge people based on their race, sex, and age. However, after going into the 20th century, people started to advocate for equality and diversity, along with the help of some civil rights leaders such as Martin Luther King, the world is slowly becoming more diversified and discrimination-free. But despite the efforts people make to promote equality, discrimination and inequality are still prominent in today’s world. For instance, according to Statistics Canada, in 2017, the number of hate crimes increased markedly by 47% and has remained at comparable levels in 2018 and 2019. Between 2018 and 2019, hate crimes motivated by hatred of a race or ethnicity increased by 10%, much of this increase was a result of more hate crimes targeting the Black and West Asian People. [@moreau_2021]

To find out more about discrimination, we obtained and analyzed data from Canadian General Social Survey, we are mostly interested in the type of discrimination that occurs most frequently and whether discrimination affects law enforcement. Additionally, we want to find out whether there is a correlation between discrimination and different type of crimes.

# Data
We started our analysis by using R [@citeR], dplyr[@dplyr], tidyverse[@tidyverse], here[@here]. Graphs are generated using and ggplot2[@ggplot2], kableExtra[@kableExtra], and gridExtra[@gridExtra].

## Data Source
We retrieved our data from CHASS [@chass]. Among all the available data in the Canadian General Social Survey(Victimization), we decide to choose the following subset: *DIS: Discrimination*, *DTS: Discrimination - Types of Situations experienced by respondent*, *CIP: Confidence In Police*, *DEM: Demographic Derived Variables*, *PCC: Perceptions: Criminal Courts*, *PLP: Perceptions: Local Police*, *TIP: Trust In People*, *CIR: Main Routing Module: Crime Incident Report*. The GSS program, established in 1985, conducts surveys annually across the 10 provinces. The GSS is recognized for its regular collection of cross-sectional data that allows for trend analysis, and its capacity to test and develop new concepts that address current and emerging issues. The two primary objectives of the General Social Survey are: 
a)  to gather data on social trends in order to monitor changes in the living conditions and well-being 
of Canadians over time; and 
b)  to provide immediate information on specific social policy issues of current or emerging interest.

To meet these objectives, the data collected by the GSS are made up of two components: classification and core content. Classification content (such as age, sex, education, income) helps to delineate population groups for use in the analysis of core data. Core content is designed to measure changes in society related to living conditions and well-being and to supply data to inform specific policy issues[@chass].

Data for 2014 General Social Survey (GSS) on Canadians’ Safety and Security (Victimization) was collected from January to December, 2014. The survey collected a large amount of information for each selected respondent as well as some information about each member of the respondent’s household. The overall response rate was 52.9%. The target population for the Cycle 28 GSS included all persons 15 years of age and older in Canada, excluding: 
1. Residents of the Yukon, Northwest Territories, and Nunavut; 
2. Full-time residents of institutions. 
The survey frame was created using two different components:

- Lists of telephone numbers in use (both landline and cellular) available to Statistics Canada from various sources (Telephone companies, Census of population, etc.);

- The Address Register (AR): List of all dwellings within the ten provinces.

## Methodology and Data Collection

```{r, include=FALSE, warning=FALSE}
gss_data <-
  read_csv(here::here("inputs/data/data.csv"))
glimpse(gss_data)
```

```{r, include=FALSE}
#Renaming the variables according to the code book by using the function "rename". 
#Select needed variables for analysis from the aspect of discrimination by using the function"select"
gss_data_renamed <- gss_data %>%
  rename(Sex = dis_10, Ethnicity_or_Culture = dis_15, Race_or_Skin_Colour = dis_20, Physical_Appearance = dis_25, Religion = dis_30, Sexual_Orientation = dis_35, Age = dis_40, Physical_or_Mental_Disability = dis_45, Language = dis_50, Other_Discrimination = dis_55, Victim_of_Discrimination_5_years = discrim, Bank_Store_Restaurant = dts_20, Work_Environment = dts_25, Police = dts_30, Courts = dts_35, Canadian_Border = dts_60, Other_Experience = dts_65) %>%
  select (Sex, Ethnicity_or_Culture, Race_or_Skin_Colour, Physical_Appearance, Religion, Sexual_Orientation, Age, Physical_or_Mental_Disability, Language, Other_Discrimination, Victim_of_Discrimination_5_years, Bank_Store_Restaurant, Work_Environment, Police, Courts, Canadian_Border, Other_Experience)

colnames(gss_data_renamed)
class(gss_data_renamed$Ethnicity_or_Culture)
```

```{r, include=FALSE}
  data_cleaned <-  gss_data_renamed %>%
  mutate( Discrimination_Reason =
           case_when(Sex == 1 ~ "Sex",
                     Ethnicity_or_Culture == 1 ~ "Ethnicity",
                     Race_or_Skin_Colour == 1 ~ "Race",
                     Physical_Appearance == 1 ~ "Physical Appearance",
                     Religion == 1 ~ "Religion",
                     Sexual_Orientation == 1 ~ "Sexual Orientation",
                     Age == 1 ~ "Age",
                     Physical_or_Mental_Disability == 1 ~ "Disability",
                     Language == 1 ~ "Language",
                     Other_Discrimination == 1 ~ "Other"
                     )) %>%
  mutate (Discrimination_Experience = 
            case_when(Bank_Store_Restaurant == 1 ~ "Bank or Store or Restaurant",
                      Work_Environment == 1 ~ "Work Environment",
                      Police == 1 ~ "Police",
                      Courts == 1 ~ "Courts",
                      Canadian_Border == 1 ~ "Cannadian Border",
                      Other_Experience == 1 ~ "Other Experience"

            ))%>%
  mutate(Have_Discrimination_Experience_in_5_years = Victim_of_Discrimination_5_years == 1)

```

The survey started with an entry section. The purpose of this section is to introduce the survey and select a respondent. A Household Roster is created, which collects key demographic information on each member of the household, including age, sex and marital status. Selected respondents are asked for their birth date and confirmation of their age (ANC). They are also asked to confirm their marital status (CMR) if another household member provided the roster information. Age and marital status are used to determine if certain questions are asked later in the survey. Age and date of birth are also used for certain derived variables and to validate responses where ages are involved[@chass].

When a probability sample is used, as is the case for the GSS, the principle behind estimation is that each person selected in the sample represents (in addition to himself or herself) several other persons not in the sample. Computer assisted telephone interviewing (CATI) was used to collect data for the 2014 GSS VIC. Respondents were interviewed in the official language of their choice. Proxy interviews were not permitted. In each frame, each record was assigned to a stratum within its province. A simple random sample without replacement of records was next selected in each stratum. Only specific geographies were targeted for the oversample of immigrants and youth[@chass].

Coverage of the 2014 GSS on Canadians’ Safety and Security (Victimization) targeted population by the survey frame is estimated to be more than 86% complete. All respondents in the ten provinces were interviewed by telephone. Households without telephones were therefore excluded. During collection, for the households not meeting the eligibility criteria, the interviews were terminated after an initial set of questions that established whether or not they met the criteria. For the oversample cases, an extra question was asked to determine if each person in the household was born in Canada or not. For the 2014 GSS on Canadians’ Safety and Security (Victimization), 81.5% of the telephone numbers dialed reached eligible households. An attempt was then made to conduct an interview with one randomly selected person from each household. Interviewers were instructed to make all reasonable attempts to obtain a completed interview with the randomly selected member of the household. Those who at first refused to participate were re-contacted up to two more times to explain the importance of the survey and to encourage their participation[@chass].

The target sample size (i.e. the number of respondents) for Cycle 28 on Canadians’ Safety and Security (Victimization) was 39,674, while the actual number of respondents was 33,127. Some GSS respondents were removed from the PUMF for confidentiality reasons. GSS Cycle 28 PUMF contains questionnaire responses and associated information from 33,089 respondents. For each province, minimum sample sizes were determined that would ensure certain estimates would have acceptable sampling variability at the stratum level. Once these stratum sample size targets had been met, the remaining sample was allocated to the strata in a way that balanced the need for precision of both national-level and stratum-level estimates[@chass].

One of the strengths of this survey is that it captures various potential factors that could have significant relationship with victimization. Also, it mainly focuses on gathering information within 5 years which is a more accurate approach for the respondent to retrieve a recent memory. However, this also comes with a trade-off that we could not compare the information in earlier periods. Dealing with non-responses is a common problem for every survey. In the GSS, the answers for each question provided different situations for the respondent to choose including Yes, No, Valid skip, Don’t know, Refusal, and Not stated. As a result, even though the existence of non-response could not be eliminated, the gathered data would provide a good sense of why the respondent could not offer a clear answer.

Another trade-off is that the general social survey is a relatively long to capture various information which may lead to impatience of answering and errors in responding. Typically, long questionnaires could involve a fake question to check if the respondent is reading the questions carefully and providing sincere answers. For some scenario questions, investigators like us would also like to obtain further elaborations. So we also include a supplementary survey to gather more detailed information about discriminations.

\newpage
# Results

## Classification of Discrimination

The identification of discrimination could be vague and inconsistent in different regions. According to Human rights in Ontario[@humanrights], there are four forms of discrimination: direct, indirect, subtle, and adverse effect discrimination; Harassment; Poisoned environment; Systemic discrimination. Discrimination often derives from different backgrounds and identities. Based on unique combination of identities, people may be exposed to specific forms of discrimination and may experience significant personal harm. The 2014 GSS on Canadians’ Safety and Security (Victimization) mainly captured gender, disability, and ethnicity. It also provides information of the condition when discrimination happens. According to Figure \@ref(fig:discrimination-reason), the most common reason of discrimination within 5 years is sex and ethnicity. The ground of “sex” is not specifically defined, although it is generally considered to be related to a person’s biological sex, male or female. Men and women receive equal protection under this ground. The ground of “sex” also includes a broader notion of “gender,” which can be described as the social characteristics attributed to each sex[@humanrights]. In addition, physical appearance, age and language are also typical reasons of discrimination.

```{r, discrimination-reason, fig.cap="Discrimination Reason", echo=FALSE, warning=FALSE}
ggplot(data = data_cleaned) + 
  geom_histogram(aes(x = Have_Discrimination_Experience_in_5_years, fill = Discrimination_Reason), 
                 position = 'dodge', 
                 bins = 10) + 
  theme(axis.text.x=element_blank()) +
  scale_x_log10()
```

Under many circumstances, the government has established regulations to protect human rights. For example, the Canadian Human Rights Act is a broad-reaching piece of legislation that prohibits discrimination on the basis of gender, race, ethnicity and other grounds. Also, Canada's Employment Equity Act and the Federal Contractors Program require employers to take active measures to improve the employment opportunities for specific groups of people in our country. Both the labour rights and responsibilities of employers and employees within federally regulated sectors fall under the Canada Labour Code. The rights of foreign workers in Canada are also protected under federal or provincial/territorial labour laws[@canada]. However, we could still see that discrimination experience is significantly common under working environment in \@ref(fig:discrimination-experience). Bank, store, and restaurant are the most common location for discrimination to take place while the Canadian border and courts are the least common location for discrimination to happen.

```{r, discrimination-experience, fig.cap="Discrimination Experience", echo=FALSE, warning=FALSE}
ggplot(data = data_cleaned) + 
  geom_histogram(aes(x = Have_Discrimination_Experience_in_5_years, fill = Discrimination_Experience), 
                 position = 'dodge', 
                 bins = 10) +
  theme(axis.text.x=element_blank()) +
  scale_x_log10()
```

## Impact of Gender on People's Confidence in Law Enforcement 
After noting that the majority of discriminating cases are gender discrimination. We are wondering if this will also be reflected in people's confidence in Law Enforcement. To be a specific, if a certain gender group are more vulnerable to discrimination, will that group also be treated unfairly during law enforcement? To investigate this, we extracted some other survey results from Canadian General Social Survey. We will first find the number of each gender in the survey, if the number of respondents from each gender groups are uneven, then the result might be biased toward a gender group. After analyzing the survey data set, the following gender distribution is found, as shown in Figure \@ref(fig:gender-distribution). From the figure, we can see that the gender distribution are about the same, so we can proceed to further investigation.

```{r, include=FALSE}
input_csv <- read_csv(here::here(
  "inputs/data/The_Impact_of_sex_on_confidence_in_government.csv"
))

gender_group <- count(input_csv, sex)
gender_group$sex[gender_group$sex==2] <- "Female"
gender_group$sex[gender_group$sex==1] <- "Male"
```

```{r, gender-distribution, fig.cap='Gender Distribution', echo=FALSE, warning=FALSE, message = FALSE, fig.width=10, fig.height=5, fig.pos = "!H"}

ggplot(data=gender_group, aes(x=sex, y=n)) +
  geom_bar(stat="identity")

```

## Confidence In Police
Now we will investigate people's confidence in police in different gender groups. People's response are categorized into 8 groups: *1  A great deal of confidence*, *2  Some confidence*, *3  Not very much confidence*, *4  No confidence at all*, *6  Valid skip*, *7  Don't know*, *8  Refusal*, *9  Not stated*. The following bar graph shows the distribution of the responses, categorized by gender groups, as shown in Figure \@ref(fig:cip-distribution). From the graph, we can see that the majority of respondents are confident in local police. To be specific, 92.8% of female show some or great confidence in local police, whereas 90% of male show some or great confidence in local police.

```{r, include= FALSE}
myvars <- c("sex", "cip_10")
cip_based_on_gender <- input_csv[myvars]
# Only interested in valid responses
cip_based_on_gender<-cip_based_on_gender[!cip_based_on_gender$cip_10>=6,]
# Rename column
cip_based_on_gender$sex[cip_based_on_gender$sex==2] <- "Female"
cip_based_on_gender$sex[cip_based_on_gender$sex==1] <- "Male"
cip_based_on_gender$cip_10[cip_based_on_gender$cip_10==1] <- "Great confidence"
cip_based_on_gender$cip_10[cip_based_on_gender$cip_10==2] <- "Some confidence"
cip_based_on_gender$cip_10[cip_based_on_gender$cip_10==3] <- "Small confidence"
cip_based_on_gender$cip_10[cip_based_on_gender$cip_10==4] <- "No confidence"
# Generate table for plotting
cip_count_based_on_gender <- table(cip_based_on_gender$sex, cip_based_on_gender$cip_10)

```

```{r, cip-distribution, fig.cap='Confidence in Police Distribution', echo=FALSE, warning=FALSE, message = FALSE, fig.width=10, fig.height=7, fig.pos = "!H"}

barplot(cip_count_based_on_gender, main="Confidence In Police Distribution",
  xlab="Confidence Level", col=c("darkblue","red"), cex.names=.8,
  legend = rownames(cip_count_based_on_gender), horiz=TRUE, beside=TRUE)
```

Next, we will take a closer look at the responses of two other questions, they are *Perception (local police) - Being approachable and easy to talk to* and *Perception (local police) - Treating people fairly*. Those questions, especially the second one, will allow us to have more certainty in determining whether people in different gender groups are treated unfairly in Law Enforcement. Both of the questions have 7 valid choices, they are *1  Good job*, *2  Average job*, *3 Poor job*, *6  Valid skip*, *7  Don't know*, *8  Refusal*, *9  Not stated*. For simplicity, we will only consider the first 3. The distribution of the responses is shown in Figure \@ref(fig:pip-distribution1) and Figure \@ref(fig:pip-distribution2). From the figures, around 81% of male respondents think that police are approachable and around 80% of female respondents hold the same opinion. On top of that, around 77% of female respondents think that police treat people fairly and around 79% of male respondents hold the same opinion. In Conclusion, based on the results of the survey, majority of the respondents from all participating gender groups have confidence in police and agree that police is approachable and treat people fairly.

```{r, include = FALSE}
myvars <- c("sex", "plp_130", "plp_140")
pip_based_on_gender <- input_csv[myvars]
# Filter out valid response
pip_based_on_gender<-pip_based_on_gender[!(pip_based_on_gender$plp_130>=6 | pip_based_on_gender$plp_140>=6),]
# Rename column
pip_based_on_gender$sex[pip_based_on_gender$sex==2] <- "Female"
pip_based_on_gender$sex[pip_based_on_gender$sex==1] <- "Male"
pip_based_on_gender$plp_130[pip_based_on_gender$plp_130==1] <- "Good Job"
pip_based_on_gender$plp_130[pip_based_on_gender$plp_130==2] <- "Average Job"
pip_based_on_gender$plp_130[pip_based_on_gender$plp_130==3] <- "Poor Job"

pip_based_on_gender$plp_140[pip_based_on_gender$plp_140==1] <- "Good Job"
pip_based_on_gender$plp_140[pip_based_on_gender$plp_140==2] <- "Average Job"
pip_based_on_gender$plp_140[pip_based_on_gender$plp_140==3] <- "Poor Job"

pip_count_based_on_gender_130 <- table(pip_based_on_gender$sex, pip_based_on_gender$plp_130)
pip_count_based_on_gender_140 <- table(pip_based_on_gender$sex, pip_based_on_gender$plp_140)

```

```{r, pip-distribution1, fig.cap='Police is approachable', echo=FALSE, warning=FALSE, message = FALSE, fig.width=10, fig.height=7, fig.pos = "!H"}

barplot(pip_count_based_on_gender_130, main="Police is Approachable and Polite",
  xlab="Confidence Level", col=c("darkblue","red"), cex.names=.8,
  legend = rownames(pip_count_based_on_gender_130), horiz=TRUE, beside=TRUE)
```

```{r, pip-distribution2, fig.cap='Police Treats People Fairly', echo=FALSE, warning=FALSE, message = FALSE, fig.width=10, fig.height=7, fig.pos = "!H"}

barplot(pip_count_based_on_gender_140, main="Police Treats People Fairly",
  xlab="Confidence Level", col=c("darkblue","red"), cex.names=.8,
  legend = rownames(pip_count_based_on_gender_140), horiz=TRUE, beside=TRUE)
```

\newpage

## Confidence In Criminal Courts
After investigating different gender group's confidence in police, we will now take a look at people's confidence in criminal courts. Similarly, we will only look at the following three responses: *1  A great deal of confidence*,
*2  Some confidence*, *3  Not very much confidence*, Figure \@ref(fig:pip-distribution3) shows the result.

```{r, include = FALSE}
myvars <- c("sex", "pcc_145")
conf_in_court_based_on_gender <- input_csv[myvars]
# Filter out valid response
conf_in_court_based_on_gender<-conf_in_court_based_on_gender[!(conf_in_court_based_on_gender$pcc_145>=6),]
conf_in_court_based_on_gender$sex[conf_in_court_based_on_gender$sex==2] <- "Female"
conf_in_court_based_on_gender$sex[conf_in_court_based_on_gender$sex==1] <- "Male"
conf_in_court_based_on_gender$pcc_145[conf_in_court_based_on_gender$pcc_145==1] <- "Great Confidence"
conf_in_court_based_on_gender$pcc_145[conf_in_court_based_on_gender$pcc_145==2] <- "Some Confidence"
conf_in_court_based_on_gender$pcc_145[conf_in_court_based_on_gender$pcc_145==3] <- "Not Very Much Confidence"
conf_in_court_based_on_gender$pcc_145[conf_in_court_based_on_gender$pcc_145==4] <- "No Confidence"
# Generate table
conf_in_court_based_on_gender <- table(conf_in_court_based_on_gender$sex, conf_in_court_based_on_gender$pcc_145)

```

```{r, pip-distribution3, fig.cap='Confidence in Criminal Courts', echo=FALSE, warning=FALSE, message = FALSE, fig.width=10, fig.height=7, fig.pos = "!H"}

barplot(pip_count_based_on_gender_140, main="Confidence in Criminal Courts",
  xlab="Confidence Level", col=c("darkblue","red"), cex.names=.8,
  legend = rownames(pip_count_based_on_gender_140), horiz=TRUE, beside=TRUE)
```

## Crime By Category

```{r crime-plot, fig.cap="Crime Count By Category", echo=FALSE, warning=FALSE, message = FALSE, fig.height=6, fig.width=8}
subset_data = gss_data %>% filter(msvic < 900)  # ignore invalid responses
subset_data %>% 
  left_join(as.data.frame(list(msvic=msvic_num, msvic_label=msvic_labels))) %>%
  ggplot(aes(as.factor(msvic_label), fill=msvic_label)) + geom_histogram(stat="count") + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(title = "Crime Count", fill="Crime Category")
```

Figure \@ref(fig:crime-plot) shows the crime count of different categories. 
We can identify the crimes that happen the most frequently.

Theft, assault, and break and enter happen most frequently.

### Crime Count By Education Level

```{r crime-count-plot-by-education, fig.cap="Crime By Education Level", echo=FALSE, warning=FALSE, message = FALSE, fig.height=6, fig.width=8}
grid.arrange(university_crime_count,
            college_crime_count,
            elementary_crime_count,
            nrow=1)

# sex_crime_plot
```

Figure \@ref(fig:crime-count-plot-by-education) show the same crime count but differentiated by education level.

The education levels are mutually exclusive (i.e. if a respondent select "Yes" for "Elementary/junior/high school" education level, then he/she can not select "Yes" for the other 2 questions). The 3 plots all have similar shapes to Figure \@ref(fig:crime-plot), i.e. the most frequent crime categories in Figure \@ref(fig:crime-plot) are also the most frequent crimes in Figure \@ref(fig:crime-count-plot-by-education).

Within Figure \@ref(fig:crime-count-plot-by-education), the subplot "University Education Crime Count" has a obviously different shape from the other 2 education level.

It seems like samples with University Education have higher chance of getting "Sexual Assault", "Vandalism" "Theft of household property".
Higher "Theft of household property" and "Vandalism" count can be easily explained by, people with higher education may have higher income (higher household property value) which could attract theft.

I couldn't find a good explanation high "Sexual Assault" for university-education population.

## Discrimination VS Crime

```{r discrim-vs-crime-plot, fig.cap="Discrimination VS Crime", echo=FALSE, warning=FALSE, message = FALSE, fig.height=6, fig.width=8}
discrim_vs_plot
```

Figure \@ref(fig:discrim-vs-crime-plot) plots the relationship between discrimination and crime to show which crimes are more related to discrimination.
Crimes like robbery and sexual assault are more likely to happen on people being discriminated.

Figure \@ref(fig:sex-crime-plot) to Figure \@ref(fig:language-crime-plot) in appendix plot the relationship between crime and different discrimination categories.

The following table shows some obvious and proportional positive relationship between crime rate and different types of discrimination (i.e. people experience discrimination $x$ is more likely to experience crime $y$).

| Discrimination                | Crime                                             |
| ----------------------------- | ------------------------------------------------- |
| Sex                           | Sexual Assault                                    |
| Ethnicity or Culture          | Robbery                                           |
| Race and Skin Color           | Robbery, Mortar theft, (Sexual) Assault           |
| Physical Appearance           | Robbery and Sexual Assault                        |
| Religion                      | Robbery and Sexual Assault                        |
| Sexual Orientation            | Attempted Robbery and Sexual Assault              |
| Age                           | Robbery and Attempted theft of household property |
| Physical or Mental Disability | Robbery and Sexual Assault                        |
| Language                      | None                                              |

\newpage

# Discussion

## Discrimination under Working Environment

Many people wrongly think that discrimination does not exist if the impact was not intended, or if there were other factors that could explain a particular situation. In fact, discrimination often takes place without any intent to do harm. And in most cases, there are overlaps between discrimination and other legitimate factors[@humanrights]. From our results, discrimination under working environment is very common. Discrimination in the workplace happens when a person or a group of people is treated unfairly or unequally because of specific characteristics. As we mentioned for discrimination reasons, these characteristics are protected. However, it still happens, even though discrimination in the workplace has been illegal for decades.

Employers have the primary obligation to make sure their workplace is free from discrimination and harassment. Employers are expected to proactively provide a workplace where human rights are respected and employees afforded equal opportunities. Even without the legal requirement not to discriminate, providing a fair work environment that respects the dignity of all employees is in an organization’s best interests. Today's global economy, multicultural society and highly competitive corporate environment place greater demands on all organizations. Employers need to recruit and retain the best employees for their organizations, and part of that process involves selecting and keeping employees with diverse backgrounds and talents from the widest possible pool of candidates.
Unions are important partners in creating a non-discriminatory workplace. Unions and employers have a joint duty to make sure that workplaces are free of discrimination and harassment. Given that around 75% of all human rights complaints arise in the workplace, it makes good business sense to take proactive action[@humanrights].

The only way to stop workplace discrimination is to speak up about these incidents, bring lawsuits, and shine a light on discriminatory practices. Most companies care about the bottom line—and they will only stop discriminating when big verdicts and settlements make discrimination too costly to permit[@king]. Another crucial reason behind the high occurrence of discrimination under working environment is that victims are usually afraid of revealing the truth since it may cost them to lose the job or being treated even more unfairly afterwards. However, to truly protect themselves away from discrimination, they should know their protected human rights and seek for help.

## Gender Discrimination in Law Enforcement

Based on results of the survey, every gender group have a similar perception in police and criminal courts. However, it is still too early to conclude that there are no gender discrimination during law enforcement because of the limitations in the survey. In the original survey, when people are asked for their gender, only Male and Female are available to choose from, the other options are either "Valid Skip" or "Refusal", as a result, every survey respondent chose either "Male" or "Female". This makes the results very biased. Suppose that all the *Poor Job* votes are from a minority gender group, the results become deceptive because they were not able to select their own gender.

## Discrimination VS Crime

One special finding is that robbery appears to be very related to discrimination.

Sexual Assault seems to be quite common in both overall crime count (Figure \@ref(fig:crime-plot)) and amoung discriminated people (Figure \@ref(fig:discrim-vs-crime-plot)),
so it's not surprise that in all types of discrimination, sex assault is almost always one of the top crime (Figure \@ref(fig:sex-crime-plot) to \@ref(fig:language-crime-plot)).

However, robbery rate has the opposite phenomenon. 
In Figure \@ref(fig:crime-plot), we can see that among all other crimes, (attempted) robbery are one of the least frequent crime.
But in Figure \@ref(fig:discrim-vs-crime-plot), robbery and attempted robbery are each as frequent as sexual assault. 
For all types of discriminations, robbery is also one of the top crime(Figure \@ref(fig:sex-crime-plot) to \@ref(fig:language-crime-plot)).

We can draw an hypothesis that people experienced discrimination are more likely to experience robbery.

\newpage

# Appendix

```{r sex-crime-plot, fig.cap="Sex Discrimination VS Crime", echo=FALSE, warning=FALSE, message = FALSE, fig.height=6, fig.width=8}
sex_crime_plot
```

```{r culture-crime-plot, fig.cap="Ethnicity or culture Discrimination VS Crime", echo=FALSE, warning=FALSE, message = FALSE}
culture_crime_plot
```

```{r race-skin-crime-plot, fig.cap="Race and Skin Color Discrimination VS Crime", echo=FALSE, warning=FALSE, message = FALSE}
race_skin_crime_plot
```

```{r physical-appearance-crime-plot, fig.cap="Physical appearance Discrimination VS Crime", echo=FALSE, warning=FALSE, message = FALSE}
physical_appearance_crime_plot
```

```{r religion-crime-plot, fig.cap="Religion Discrimination VS Crime", echo=FALSE, warning=FALSE, message = FALSE}
religion_crime_plot
```

```{r sexual-orientation-crime-plot, fig.cap="Sexual orientation Discrimination VS Crime", echo=FALSE, warning=FALSE, message = FALSE}
sexual_orientation_crime_plot
```

```{r age-crime-plot, fig.cap="Age Discrimination VS Crime", echo=FALSE, warning=FALSE, message = FALSE}
age_crime_plot
```

```{r physical-mental-disability-crime-plot, fig.cap="Physical or mental disability Discrimination VS Crime", echo=FALSE, warning=FALSE, message = FALSE}
physical_mental_disability_crime_plot
```

```{r language-crime-plot, fig.cap="Language Discrimination VS Crime", echo=FALSE, warning=FALSE, message = FALSE}
language_crime_plot
```

```{r fig.width=8, fig.height=5, echo=FALSE, warning=FALSE, message = FALSE}
cross_compare_barplot(data=gss_data %>% filter(msvic < 900) %>% filter(eduf10 < 6),
                            col1 = "edum5", col2 = "msvic",
                      col1_num=eduf10_num, col1_labels=eduf10_labels,
                      col2_num=msvic_num, col2_labels=msvic_labels, 
                      plot_title="Father Education VS Crime", 
                      plot_subtitle="Relatively low crime rate for all crimes", 
                      plot_xlab="Crime", 
                      plot_ylab="Percentage of Father Education", 
                      plot_fill_lab="Education")
```

\newpage

# Reference